---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# trdizinR

<!-- badges: start -->
<!-- badges: end -->

The **trdizinR** package provides a convenient R interface to access the [TR Dizin](https://trdizin.gov.tr/) API. It allows users to programmatically search and retrieve metadata on academic **publications**, **authors**, **journals**, and **institutions** indexed by TR Dizin.

## Installation

```{r, eval=FALSE}
# Install from GitHub
remotes::install_github("emraher/trdizinR")
```


## API Overview

TR Dizin provides access to the following data:

- **Publications**
- **Journals**
- **Authors**
- **Institutions**

This package wraps the relevant endpoints and parses the JSON responses into R-readable list objects.

## Functions

| Function                   | Description                                 |
|----------------------------|---------------------------------------------|
| `search_publication()`     | Search publications                         |
| `search_journal()`         | Search journals                             |
| `search_author()`          | Search authors                              |
| `search_institution()`     | Search institutions                         |
| `get_publication_detail()` | Get details for a specific publication      |
| `get_journal_detail()`     | Get details for a specific journal          |
| `get_author_detail()`      | Get details for a specific author           |
| `get_institution_detail()` | Get details for a specific institution      |

## Usage

```{r}
library(trdizinR)

# Search for publications
result <- search_publication("cancer")

# Search for journals
result_journal <- search_journal("cancer")

# Search for authors
result_author <- search_author("Civcir")

# Search for institutions
result_institution <- search_institution("University")

# Get details of a publication
res <- search_publication("medicine", limit = 1)
publication_id <- res$hits$hits$`_id`[1]
publication_detail <- get_publication_detail(publication_id)


# Get details of a journal
res <- search_journal("medicine", limit = 1)
journal_id <- res$hits$hits$`_id`[1]
journal_detail <- get_journal_detail(journal_id)


# Get details of an author
res <- search_author("Sunay Çelik", limit = 1)
author_id <- res$hits$hits$`_id`[1]
author_detail <- get_author_detail(author_id)


# Get details of an institution
res <- search_institution("mudanya", limit = 1)
inst_id <- res$hits$hits$`_source`$codes[1]
inst_detail <- get_institution_detail(inst_id)


```

## Faceted Search Support

All `search_*()` functions in the `trdizinR` package (e.g., `search_publication()`, `search_journal()`, `search_author()`, `search_institution()`) support **faceted filtering** through additional named arguments using the `...` ellipsis operator.

### What Are Facets?

**Facets** are filters that narrow search results by specific metadata fields. They correspond to fields like year, institution type, country, author ORCID, and more—depending on the type of entity being searched.

These facets align with the [TRDizin API documentation](https://development.trdizin.gov.tr/#tanml-kok-adresleri), which describes how to apply these filters to the base query.

### Example Usage

```{r}
# Search publications related to cancer, published in 2020
res <- search_publication("cancer", facet_publication_year = 2020)

# Search journals with a specific year filter
res <- search_journal("education", facet_year = 2023)

# Search institutions filtered by country
res <- search_institution("University", facet_country = "Türkiye")

# Search authors filtered by minimum publication count
res <- search_author("Civcir", facet_minPublicationCount = 10)
```


### Supported Facets 

This section provides a comprehensive list of supported facet parameters for the TR Dizin API, organized by entity type. These facets can be used to filter search results in `trdizinR` package functions.

#### Publication Facets

| Facet Parameter                  | Description                                         | Example Values                                                                 |
|----------------------------------|-----------------------------------------------------|---------------------------------------------------------------------------------|
| `facet_documentType`             | Filters by document type                            | `PAPER`, `PROJECT`                                                              |
| `facet_database`                 | Filters by database type                            | `SOCIAL`, `SCIENCE`                                                             |
| `facet_accessType`              | Filters by access type                              | `OPEN`, `CLOSE`                                                                 |
| `facet_authorName`              | Filters by author name                              | `Ali Yıldız`, `Sunay Çelik`                                                     |
| `facet_publication_year`         | Filters by publication year                         | `2020`, `2021`                                                                  |
| `facet_publicationLanguage`      | Filters by publication language                     | `ENG`, `TUR`, `GER`, `FRE`, `ARA`, `OTH`, `RUS`, `ITA`, `SPA`                   |
| `facet_journalName`              | Filters by journal name                             | `Milli Eğitim`, `Medicine Science`                                              |
| `facet_publicationType`          | Filters by publication type                         | `RESEARCH`, `COMPILATION`, `FACT_PRESENTATION`, `OTHER`, `BOOK_PRESENTATION`, `CORRECTION`, `EDITORIAL`, `LETTER`, `LETTER_TO_EDITOR`, `MEETING_SUMMARY`, `REPORT`, `SHORT_REPORT`, `TRANSLATION`, `RETRACTED` |
| `facet_projectGroup`             | Filters by project group                            | `TÜBİTAK ÇAG Proje`                                                             |
| `facet_facetauthorInstitution`   | Filters by author's institution                     | `T.C. MİLLİ EĞİTİM BAKANLIĞI`                                                   |
| `facet_facetauthorCountry`       | Filters by author's country                         | `Türkiye`, `İtalya`                                                             |
| `facet_facetauthorCity`          | Filters by author's city                            | `ANKARA`, `KONYA`                                                               |
| `facet_subject`                  | Filters by subject area                             | `Fen > Temel Bilimler > Biyoloji`, `Sosyal > Sosyal > Psikoloji`                |

#### Journal Facets

| Facet Parameter             | Description                             | Example Values                                                                 |
|-----------------------------|-----------------------------------------|---------------------------------------------------------------------------------|
| `facet_year`                | Filters by year                         | `2020`, `2021`                                                                  |
| `facet_publicationLanguage` | Filters by publication language         | `ENG`, `TUR`, `GER`, `FRE`, `ARA`, `OTH`, `RUS`, `ITA`, `SPA`                   |
| `facet_subject`             | Filters by subject area                 | `Fen > Temel Bilimler > Biyoloji`, `Sosyal > Sosyal > Psikoloji`                |
| `facet_authorStatus`        | Filters by author status                | `REAL`, `POOL`                                                                  |

#### Author Facets

| Facet Parameter             | Description                             | Example Values                                                                 |
|-----------------------------|-----------------------------------------|---------------------------------------------------------------------------------|
| `facet_orcid`               | Filters by ORCID                        | `0000-0002-1825-0097`                                                           |
| `facet_minPublicationCount` | Filters by minimum publication count    | `10`, `20`                                                                      |

#### Institution Facets

| Facet Parameter     | Description                             | Example Values                                                                 |
|---------------------|-----------------------------------------|---------------------------------------------------------------------------------|
| `facet_type`        | Filters by institution type             | `DEVLETUNIVERSITE`, `VAKIFUNIVERSITE`, `KAMUKURULUS`, `DIGER`, `KURUMKURULUS`, `ASKERI`, `VOD` |
| `facet_status`      | Filters by institution status           | `OZEL`, `KAMU`, `DIGER`, `YABANCI`, `STK`                                       |
| `facet_country`     | Filters by country                      | `Türkiye`, `Brezilya`                                                           |

## Notes

- All functions use `httr` and `jsonlite` to make and parse API calls.
- Additional facet filters can be passed using the `...` argument in the search functions.
- When using these facets with `search_*()` functions in R, pass them as named arguments with the required `facet_` prefix.

## License

MIT © [Emrah Er](https://github.com/emraher)
